Accelerated Magnetic Resonance Imaging by Adversarial Neural Network

### 2018-02-17 paper.tex
English editing

We introduce a practical, software-only framework, based on deep learning, for accelerating MRI acquisition, while maintaining anatomically meaningful imaging. This is accomplished by partial MRI sampling, while using an adversarial neural network to directly estimate the missing k-space samples. The inter-play between the generator and the discriminator networks enables the introduction of an adversarial cost in addition to a fidelity loss used for optimizing the peak signal-to-noise ratio (PSNR). Promising image reconstruction results are obtained for 3T and 1.5T brain MRI, from large publicly available dataset, where only 40\%, 25\% and 16.6\% of the raw samples of each scan are used. To assess the clinical usability of the reconstructed images we also performed tissue segmentation and compared the results to those obtained by using the original fully-sampled images.
Segmentation compatibility, measured in terms of Dice scores and Hausdorff distances, demonstrate the quality of the proposed MRI reconstruction with respect to other methods, including the widely-used Compressed Sensing.

@@

\section{Introduction}
Magnetic Resonance Imaging (MRI) is a non-ionizing imaging modality, and is therefore widely used in diagnostic medicine and biomedical research. The physical principles of MRI are based on a strong magnetic field and pulses of radio frequency (RF) electromagnetic radiation. 
Images are produced when hydrogen atoms, which are prevalent in living organisms, emit the absorbed RF energy that is then received by antennas in close proximity to the anatomy being examined. Spatial localization of the detected MRI signals is obtained by varying the magnetic field gradients. The discretized RF output is presented in a Fourier space
(called K-space), where the x-axis is refers to the frequency and the y-axis to the phase. An inverse fast Fourier transform (IFFT) of the K-space is then used for generating anatomically meaningful MRI scans. Figure \ref{fig:pair_of_k_space_mri} presents K-space traversal patterns used in conventional imaging. Each row of the k-space is acquired after one RF excitation pulse. The number of rows multiplied by the number of slices (z-axis) determines the total scan time.

The duration of standard single structural MRI acquisition is approximately 5 minutes. Usually, several scans of different modalities or a sequence of scans are acquired such that the overall scan time is much longer. Lengthy imaging process reduces patient comfort and is more vulnerable to motion artifacts. In cases where motion is inevitable, e.g., fetal MRI, cardiac cine, and lungs imaging, scan time must be significantly shortened, otherwise the produced images might be useless. Moreover, in dynamic MRI sequences, acquisition must be brief such that the temporal resolution of the sequence would allow capturing significant temporal changes, e.g., instantaneous increment of the contrast-enhanced material concentration in DCE-MRI or differences in hemodynamic response expressed in fMRI \cite{moeller2010multiband}.

A straight forward reduction of the scan time can be obtained by sampling fewer slices, thus reducing the spatial resolution in the z-axis. Spatial distances between adjacent slices of fetal MRI or fMRI, for example, are often as high as 0.5 centimeters. Therefore, a significant portion of the potential input is not covered through imaging. On the other hand, under-sampling in the x-y domain leads to aliasing, as predicted by the Nyquist sampling theorem.

@@

The compressed sensing (CS) technique~\cite{donoho2006compressed} allows efficient acquisition and reconstruction of a signal with fewer samples than the Nyquist-Shannon sampling theorem requires, if the signal has sparse representation in a known transform domain. Using CS for MRI reconstruction by sampling a small subset of the k-space grid had been proposed in~\cite{lustig2007sparse}. The underlying assumption is that the undersampling is random, such that the zero-filled Fourier reconstruction exhibits incoherent artifacts that behave similarly to additive random noise. 
The smooth reconstruction obtained by the CS, may result in a loss of noise-like texture that can be anatomically meaningful.

@@

Deep neural networks (DNN) allow to implicitly learn prior information from large datasets. This has been exploited in recent works for MRI reconstruction from sub-sampled data. A convolutional neural network (CNN) was introduced in~\cite{wang2016accelerating} to learn the mapping between zero-filled and fully-sampled MRI.
In~\cite{Oktay2016}, MRI acceleration by super-resolution (SR) concepts was introduced. One of the key ideas is a siamese CNN structure that allows to receive multiple inputs acquired from different viewing planes. Improved low-resolution (LR) high-resolution (HR) mapping was obtained by residual architecture.
A multi-scale residual network, also known as U-net~\cite{ronneberger2015u}, was used in~\cite{lee2017deep} to address the CS optimization, formulated as a residual regression problem.
Hayun at el.~\cite{hyun2017deep}, applied U-net to perform anti-aliasing of MRI generated from sub-sampled k-space. The U-net output image is than enhanced by integrating its k-space domain with the originally acquired raw samples. U-net CNN have been also used in~\cite{sandinodeep} for dynamic MRI reconstruction.
The spatio-temporal characteristics of sequential MRI data was recently exploited using recurrent neural network (RNN) by~\cite{qin2017convolutional}. All these works address the reconstruction problem in the image domain rather than the k-space domain.

The proposed framework utilizes recent advances in deep learning, while addressing MRI reconstruction directly from the k-space. Specifically, we use generative adversarial networks (GAN) \cite{goodfellow2014generative,radford2015unsupervised,pathak2016context}.
GANs are based on the inter-play between two networks: a generator and a discriminator. The generator is capable of learning the distribution over a data-base, and sample realizations of this distribution. 
The discriminator is trained to distinguish between `generated' samples and real ones. This powerful combination has been used for MRI acceleration in~\cite{yu2017deep,mardani2017deep}, and also for generating Computed Tomography (CT)-like images from MRIs~\cite{nie2016medical}.
In contrast, here, the generator is used for reconstruction of the entire k-space grid from under-sampled data. Its loss is a combination of an adversarial loss, calculated on the image domain, based on the discriminator output, and a fidelity loss with respect to the fully sampled k-space. Unlike~\cite{hyun2017deep}, the entire process, including the IFFT, is carried out and optimized in an end-to-end manner by the proposed GAN architecture. As the MRI data is complex, both real and imaginary components are incorporated in both the generator and the discriminator.

We compare the proposed method reconstructions to those obtained by using a conventional compressed sensing method CS-MRI \cite{lustig2007sparse} and Zero-filling. In addition, for comparison, we test two other networks: 1. CNN-L2 - a k-space generator (with same architecture as the proposed method) trained with only L2 loss. 2. IM-CNN-L2 - a network which performs on the image-space and is optimized to remove the artifacts caused by under-sampling and zero padding.
The method is applied to the Information eXtraction from Images (IXI) dataset, which is a large publicly available dataset, including 3T and 1.5T scans. Raw data is down-sampled such that only 40\%, 25\% and 16.6\% of the raw samples of each scan are used.
To assess the clinical usability of the reconstructed images we perform an extended validation, composed of anatomical quantitative measurements. In addition to the commonly used PSNR and Structure Similarity Index (SSIM) we perform tissue segmentation and compare the results to those obtained by using the original fully-sampled images.
Segmentation compatibility, measured in terms of Dice scores and Hausdorff distances, demonstrate the quality of the proposed MRI reconstruction with respect to other methods.
The proposed method outperforms all other tested methods in all evaluation metrics. Moreover, compared to the widely used CS-MRI, the proposed method is faster by two orders of magnitude. In fact, it can work in real time, and is therefore practical for routine uses.

The rest of the paper is organized as follows: Section~\ref{method_section} provides a brief background on MRI acquisition and introduce the network architecture as well as the training methodology. Section~\ref{experiments_section} presents the experimental results. Conclusions and future directions are given in section~\ref{conclusions_section}.

%--------------------- METHOD --------------------%
\section{Method}\label{method_section}
We aim to address MRI reconstruction from under-sampled k-space data using DNN. We hereby present the proposed methodology as well as the network architecture.
\subsection{Notations and Problem Definition}
Let $\bm{u}$ denote the desired signal, a 2D MR image, obtained by IFFT of the complex k-space signal $s_{0}$ i.e, 
\begin{equation}
\bm{u}=F^{H}s_{0}
\end{equation} 
where $H$ is the Hermitian transpose operation and $F^{H}$ is an orthonormal 2D IFFT operator, such that $F^{H}F=I$. 
Partial k-space sampling, results in MR images that are subject to artifacts and aliasing. 
Figure~\textbf{\ref{fig:pair_of_k_space_mri}}.d exemplifies artifacts caused by under-sampling and zero padding in the phase axis using a Gaussian mask (40\%).

@@

Let $s_{p}=M_{p}\odot s_{0}$ denote the under-sampled k-space, where $\odot$ denotes element-wise multiplication and $M_{p}$ is the sampling mask. Given a model $f$, defined by the set of parameters $\Theta$, our goal is to estimate the missing k-space samples such that: 
\begin{equation}
\Theta=\arg\min_{\Theta}L(F^{H}f\left(s_{p};\Theta\right),\boldsymbol{u})
\end{equation}
where $L\left(\cdot\right)$ is a loss function. L2 loss is widely used for natural image reconstruction. However, as mentioned in \cite{pathak2016context}, L2 minimization provides a blurry solution, results in a loss of fine details. In our case, where spatial frequencies in the form of k-space are considered, using this loss alone may be inappropriate, resulting in poor reconstruction. To address this problem, we propose using in addition an adversarial loss, based on a GAN.

@@

\begin{equation}
\min_{G}\max_{D}V(G,D) = \mathbb{E_{\mathrm{s_{0}\sim p_{r}\left(s_{0}\right)}}\mathrm{\log\left[D\left(s_{0}\right)\right]}}+\mathbb{E_{\mathrm{z\sim p_{u}\left(z\right)}}\mathrm{\log\left[1-D\left(G(z)\right)\right]}}
\end{equation}
where, $z$ is a noise drawn from a uniform distribution $p_{u}\left(z\right)$.

@@

\begin{equation}
\begin{aligned}L_{G}=\alpha\cdot E_{\mathrm{z\sim p_{u}\left(z\right)}}\mathrm{\log\left[1-D\left(F^{-1}\left(\hat{s}_{0}\right)\right)\right]}+\\
\beta\cdot||\left(1-M_{p}\right)\odot\left(\hat{s}_{0}-s_{0}\right)||_{2}^{2}
\end{aligned}
\end{equation}
where, $\hat{s}_{0}$ is the estimated k-space and $\alpha=1,\,\beta=1$ are hyperparameters tuned by a cross-validation process.
We choose to use the Wasserstein distance as the GAN loss function (WGAN)~\cite{arjovsky2017wasserstein}, which improves the training stability and provides better reconstruction compared to the classic GAN loss we used in our preliminary work~\cite{shitrit2017accelerated}.

@@

The training data consists of 500 $3D$ brain MRI (T1-weighed) scans of different patients from the IXI dataset\footnote{\url{http://brain-development.org/ixi-dataset/}}. The data was acquired by three MR machines, Philips 1.5T, 3T and GE 1.5T using different imaging protocols.
From each $3D$ volume we extract 93 $2D$ saggital slices. All slices were interpolated to $256\times256$ pixels. We used $37.2k$ $(80\%)$ 2D slices for training and $9.3k$ $(20\%)$ for testing (100 $3D$ volumes).
In order to create k-space images for training, inverse orthonormal
2D FFT was applied to the fully-sampled MR images. 
Unlike our preliminary work~\cite{shitrit2017accelerated}, which employed 1D Cartesian sampling mask along the phase axis, here, we perform k-space sampling using $2D$ random Gaussian mask. By this, we exploit the sparsity in both phase and frequency dimensions and are therefore able to accommodate higher sampling factors (Figure~\ref{fig:sampling_masks}). MRI reconstruction is performed using 40\%, 25\% and 16.6\% of the original k-space data.
%Data augmentation is created by random offsets of raw images (prior to the IFFT) and k-space flipping. This leads us to reconstruction of the MR image from 40\%, 25\% and 16.6\% of the original k-space data.

The generator is composed of $5$ blocks of CONV-BatchNorm-ReLU, with output channels $16,32,64,32,8,$ respectively (see Figure~\ref{fig:architectures}). The last CONV layer is composed of two output channels for real and imaginary components. The discriminator is composed of $4$ blocks of CONV-Pool-BatchNorm-LReLU with output channels $8,16,32,16$ and one fully-connected layer. All CONV layer's kernel size is $3\times3$. All weights were initialized using Xavier \cite{glorot2010understanding}. We used the RMSprop~\cite{tieleman2012lecture} solver with a fixed learning rate of $5\mathrm{e}{-6}$ and set $k_{d}$ to 1.

We compared the proposed method reconstructions to those obtained by using a conventional compressed sensing method-CS-MRI \cite{lustig2007sparse} and Zero-filling. In addition, we tested two other networks. To asses the use of an adversarial loss, we trained a k-space generator, with the same architecture as the proposed method, with only a L2 loss (CNN-L2). We also implemented the method proposed in~\cite{hyun2017deep} and constructed a network IM-CNN-L2, which is applied directly to the image space rather than the k-space. Specifically, as in~\cite{hyun2017deep}, we used the U-net architecture~\cite{ronneberger2015u} and applied an FFT on each of its output images to get an estimated k-space. The complete k-space is then recovered by integrating the estimated samples with the sampled raw data according to the sampling mask. The same sampling masks were used for all methods.

To demonstrate the clinical usability of the reconstructed images we perform an extended validation, composed of anatomical quantitative measurements. In addition to the common used PSNR and Structure Similarity Index (SSIM)~\cite{wang2004image} metrics, we perform tissue segmentation and compared the results to those obtained by using the original fully-sampled images. The segmentation compatibility is measured in terms of Dice scores and Hausdorff distances.

@@

\subsection{PSNR and SSIM}
PSNR and SSIM are the most commonly used measures for evaluation of image reconstruction quality. PSNR measures the Mean Squared Error (MSE) between the fully-sampled MR image and the reconstructed image.
Quantitative evaluation is presented in~Table~\ref{tbl:PSNR_NO_MASK} and a graphical visualization in ~Figure~\ref{fig:error_psnr_errorbar}. The PSNR is calculated for the entire image without masking. Note that the proposed method outperforms the others. To support this, we also calculated the Two-sample t-Test for the table's scores. The differences are statistically significant (p-values~$<0.005$) for all pairwise comparisons between the results of our method and the others.

@@

An additional validation of our model is performed by calculating the PSNR measure on different brain tissues. We applied the commonly used FAST~\cite{zhang2001segmentation} image segmentation algorithm, on the original fully-sampled MR image and then used it for calculating the masked-PSNR on gray matter (GM), white matter (WM) and the Cerebrospinal Fluid (CSF). Results are presented in~Table~\ref{tbl:PSNR_WITH_MASK}. 
The differences are statistically significant with p-values~$<0.0005$ for CS-MRI and Zero-filled for sampling factors 2.5 and 4, and also for IM-CNN-L2 with p-values~$<0.0001$ for sampling factor 6.

@@

We next used SSIM, which is a full reference metric, and is considered as a perception-based model, to measure the similarity between the original fully-sampled images and the reconstructed ones.
It quantifies the degradation caused by sampling artifacts based on pixels' inter-dependencies in the image. 
%The SSIM between images $I_{\mbox{\tiny{REF}}}$ and $I_{\mbox{\tiny{REC}}}$ is calculated on a multiples sliding windows with size of $N\times N$ as follows:
%\begin{equation}
%\begin{array}{cc}
%  SSIM(I_{\mbox{\tiny{REF}}},I_{\mbox{\tiny{REC}}}) = \frac{(2\mu_{I_{\mbox{\tiny{REF}}}}\mu_{I_{\mbox{\tiny{REC}}}} + c_1) (2\sigma_{I_{\mbox{\tiny{REF,REC}}}} + c_2)}{(\mu_{I_{\mbox{\tiny{REF}}}}^2 + \mu_{I_{\mbox{\tiny{REC}}}}^2 + c_1) (\sigma_{I_{\mbox{\tiny{REF}}}}^2) + \sigma_{I_{\mbox{\tiny{REC}}}}^2) + c_2} \\
% second = 2\\
% ohad
%\end{array}
%\end{equation}
%Let $\mu_1,\mu_2$ denote the brain boundaries ext
SSIM scores for different sampling factors, for all methods compared, are presented in~Table~\ref{tbl:SSIM_NO_MASK}. The advantage of our method is clearly shown. All pairwise comparisons between the results obtained by our method and the others are statistically significant (p-values~$<0.0001$) for all sampling factors.

@@

Let $C_1,C_2\in R^2$ denote the brain boundaries extracted by the BET algorithm, using the original and the reconstructed scans respectively. The MHD between $C_1$ and $C_2$ is:

\begin{equation}
\begin{array}{cc}
MHD(C_1,C_2) = 0.5 \left[{\frac{1}{|C_1|} \sum_{\sigma_{1}\in C_1}^{}d(\sigma_1,C_2) + \frac{1}{|C_2|} \sum_{\sigma_{2}\in C_2}^{}d(\sigma_2,C_1)}\right] \\
d(\sigma_1,C_2) = \underset{\sigma_{2}\in C_{2}}{\min}||\sigma_1-\sigma_2|| \\
d(\sigma_2,C_1) = \underset{\sigma_{1}\in C_{1}}{\min}||\sigma_2-\sigma_1||
\end{array}
\end{equation}
where $\sigma_1,\sigma_2$ are points on the contours $C_1,C_2$ respectively (Figure~\ref{fig:brain_extraction}.c) and $|C_i|,~i=1,2$ are the respective cardinalities. MHD values for all tested methods and sampling ratios are presented in~Table~\ref{tbl:MHD}. The MHD results obtained by our methods are better than all others and are significant (p-values~$<0.0001$) for CS-MRI and Zero-filled.

@@

	\caption{\textcolor{black}{\footnotesize{}{}Segmentation Dice score for different sampling ratios}{\footnotesize{}\label{tbl:DICE}}}

>>>>

We introduce a practical, software-only framework, based on deep learning, for accelerating MRI acquisition, while maintaining anatomically meaningful imaging. This is accomplished by partial MRI sampling, while using an adversarial neural network to directly estimate the missing k-space samples. The inter-play between the generator and the discriminator networks enables the introduction of an adversarial cost in addition to fidelity loss used for optimizing the peak signal-to-noise ratio (PSNR). Promising image reconstruction results are obtained for 3T and 1.5T brain MRI, from a large publicly available dataset, where only 40\%, 25\%, and 16.6\% of the raw samples of each scan are used. To assess the clinical usability of the reconstructed images we also performed tissue segmentation and compared the results to those obtained by using the original fully-sampled images.
Segmentation compatibility, measured in terms of Dice scores and Hausdorff distances, demonstrate the quality of the proposed MRI reconstruction with respect to other methods, including the widely-used Compressed Sensing.

@@

\section{Introduction}
Magnetic Resonance Imaging (MRI) is a non-ionizing imaging modality, and is therefore widely used in diagnostic medicine and biomedical research. The physical principles of MRI are based on a strong magnetic field and pulses of radio frequency (RF) electromagnetic radiation. 
Images are produced when hydrogen atoms, which are prevalent in living organisms, emit the absorbed RF energy that is then received by antennas in close proximity to the anatomy being examined. Spatial localization of the detected MRI signals is obtained by varying the magnetic field gradients. The discretized RF output is presented in a Fourier space
(called k-space), where the x-axis refers to the frequency and the y-axis to the phase. An inverse fast Fourier transform (IFFT) of the k-space is then used for generating anatomically meaningful MRI scans. Figure \ref{fig:pair_of_k_space_mri} presents k-space traversal patterns used in conventional imaging. Each row of the k-space is acquired after one RF excitation pulse. The number of rows multiplied by the number of slices (z-axis) determines the total scan time.

The duration of standard single structural MRI acquisition is approximately 5 minutes. Usually, several scans of different modalities or a sequence of scans are acquired such that the overall scan time is much longer. The lengthy imaging process reduces patient comfort and is more vulnerable to motion artifacts. In cases where motion is inevitable, e.g., fetal MRI, cardiac cine, and lungs imaging, scan time must be significantly shortened, otherwise the produced images might be useless. Moreover, in dynamic MRI sequences, acquisition must be brief such that the temporal resolution of the sequence would allow capturing significant temporal changes, e.g., instantaneous increment of the contrast-enhanced material concentration in DCE-MRI or differences in hemodynamic response expressed in fMRI \cite{moeller2010multiband}.

A straight forward reduction of the scan time can be obtained by sampling fewer slices, thus reducing the spatial resolution in the z-axis. Spatial distances between adjacent slices of fetal MRI or fMRI, for example, are often as large as 0.5 centimeters. Therefore, a significant portion of the potential input is not covered through imaging. On the other hand, under-sampling in the x-y domain leads to aliasing, as predicted by the Nyquist sampling theorem.

@@

The compressed sensing (CS) technique~\cite{donoho2006compressed} allows efficient acquisition and reconstruction of a signal with fewer samples than the Nyquist-Shannon sampling theorem requires, if the signal has sparse representation in a known transform domain. Using CS for MRI reconstruction by sampling a small subset of the k-space grid had been proposed in~\cite{lustig2007sparse}. The underlying assumption is that the undersampling is random, such that the zero-filled Fourier reconstruction exhibits incoherent artifacts that behave similarly to additive random noise. 
The smooth reconstruction obtained by the CS may result in a loss of noise-like texture that can be anatomically meaningful.

@@

Deep neural networks (DNN) allow implicitly learning prior information from large datasets. This has been exploited in recent works for MRI reconstruction from sub-sampled data. A convolutional neural network (CNN) was introduced in~\cite{wang2016accelerating} to learn the mapping between zero-filled and fully-sampled MRIs.
In~\cite{Oktay2016}, MRI acceleration by super-resolution (SR) concepts was introduced. One of the key ideas is a siamese CNN structure that allows receiving multiple inputs acquired from different viewing planes. Improved low-resolution (LR) high-resolution (HR) mapping was obtained by residual architecture.
A multi-scale residual network, also known as U-net~\cite{ronneberger2015u}, was used in~\cite{lee2017deep} to address the CS optimization, formulated as a residual regression problem.
Hayun et el.~\cite{hyun2017deep}, applied U-net to perform anti-aliasing of an MRI generated from sub-sampled k-space. The U-net output image is then enhanced by integrating its k-space domain with the originally acquired raw samples. U-net CNN have been also used in~\cite{sandinodeep} for dynamic MRI reconstruction.
The spatio-temporal characteristics of sequential MRI data was recently exploited using recurrent neural network (RNN) by~\cite{qin2017convolutional}. All these works address the reconstruction problem in the image domain rather than the k-space domain.

The proposed framework utilizes recent advances in deep learning, while addressing MRI reconstruction directly from the k-space. Specifically, we use generative adversarial networks (GAN) \cite{goodfellow2014generative,radford2015unsupervised,pathak2016context}.
GANs are based on the inter-play between two networks: a generator and a discriminator. The generator is capable of learning the distribution over a data-base, and sample realizations of this distribution. 
The discriminator is trained to distinguish between `generated' samples and real ones. This powerful combination has been used for MRI acceleration in~\cite{yu2017deep,mardani2017deep}, and also for generating Computed Tomography (CT)-like images from MRIs~\cite{nie2016medical}.
In contrast, here, the generator is used for reconstruction of the entire k-space grid from under-sampled data. Its loss is a combination of an adversarial loss, calculated on the image domain based on the discriminator output, and a fidelity loss with respect to the fully sampled k-space. Unlike~\cite{hyun2017deep}, the entire process, including the IFFT, is carried out and optimized in an end-to-end manner by the proposed GAN architecture. As the MRI data is complex, both real and imaginary components are incorporated in both the generator and the discriminator.

We compare the proposed method reconstructions to those obtained by using a conventional compressed sensing method CS-MRI \cite{lustig2007sparse} and zero-filling. In addition, for comparison, we tested two other networks: 1. CNN-L2 - a k-space generator (with same architecture as the proposed method) trained with only L2 loss; 2. IM-CNN-L2 - a network that performs on the image-space and is optimized to remove the artifacts caused by under-sampling and zero padding.
The method is applied to the Information eXtraction from Images (IXI) dataset, which is a large publicly available dataset, including 3T and 1.5T scans. Raw data is down-sampled such that only 40\%, 25\%, and 16.6\% of the raw samples of each scan are used.
To assess the clinical usability of the reconstructed images we perform an extended validation, composed of anatomical quantitative measurements. In addition to the commonly used PSNR and Structure Similarity Index (SSIM), we perform tissue segmentation and compare the results to those obtained by using the original fully-sampled images.
Segmentation compatibility, measured in terms of Dice scores and Hausdorff distances, demonstrate the quality of the proposed MRI reconstruction with respect to other methods.
The proposed method outperforms all other tested methods in all evaluation metrics. Moreover, compared to the widely used CS-MRI, the proposed method is faster by two orders of magnitude. In fact, it can work in real time, and is therefore practical for routine uses.

The rest of the paper is organized as follows: Section~\ref{method_section} provides a brief background on MRI acquisition and introduces the network architecture as well as the training methodology. Section~\ref{experiments_section} presents the experimental results. Conclusions and future directions are given in section~\ref{conclusions_section}.

%--------------------- METHOD --------------------%
\section{Method}\label{method_section}
We aim to address MRI reconstruction from under-sampled k-space data using DNN. We hereby present the proposed methodology as well as the network architecture.
\subsection{Notations and Problem Definition}
Let $\bm{u}$ denote the desired signal, a 2D MR image, obtained by IFFT of the complex k-space signal $s_{0}$, i.e., 
\begin{equation}
\bm{u}=F^{H}s_{0}
\end{equation} 
where $H$ is the Hermitian transpose operation and $F^{H}$ is an orthonormal 2D IFFT operator, such that $F^{H}F=I$. 
Partial k-space sampling, results in MR images that are subject to artifacts and aliasing. 
Figure~\textbf{\ref{fig:pair_of_k_space_mri}}.d exemplifies artifacts caused by under-sampling and zero padding in the phase axis using a Gaussian mask (40\%).

@@

Let $s_{p}=M_{p}\odot s_{0}$ denote the under-sampled k-space, where $\odot$ denotes element-wise multiplication and $M_{p}$ is the sampling mask. Given a model $f$, defined by the set of parameters $\Theta$, our goal is to estimate the missing k-space samples such that: 
\begin{equation}
\Theta=\arg\min_{\Theta}L(F^{H}f\left(s_{p};\Theta\right),\boldsymbol{u})
\end{equation}
where $L\left(\cdot\right)$ is a loss function. L2 loss is widely used for natural image reconstruction. However, as mentioned in \cite{pathak2016context}, L2 minimization provides a blurry solution, resulting in a loss of fine details. In our case, where spatial frequencies in the form of k-space are considered, using this loss alone may be inappropriate, resulting in poor reconstruction. To address this problem, we propose using in addition an adversarial loss, based on a GAN.

@@

\begin{equation}
\min_{G}\max_{D}V(G,D) = \mathbb{E_{\mathrm{s_{0}\sim p_{r}\left(s_{0}\right)}}\mathrm{\log\left[D\left(s_{0}\right)\right]}}+\mathbb{E_{\mathrm{z\sim p_{u}\left(z\right)}}\mathrm{\log\left[1-D\left(G(z)\right)\right]}}
\end{equation}
where $z$ is a noise drawn from a uniform distribution $p_{u}\left(z\right)$.

@@

\begin{equation}
\begin{aligned}L_{G}=\alpha\cdot E_{\mathrm{z\sim p_{u}\left(z\right)}}\mathrm{\log\left[1-D\left(F^{-1}\left(\hat{s}_{0}\right)\right)\right]}+\\
\beta\cdot||\left(1-M_{p}\right)\odot\left(\hat{s}_{0}-s_{0}\right)||_{2}^{2}
\end{aligned}
\end{equation}
where $\hat{s}_{0}$ is the estimated k-space and $\alpha=1,\,\beta=1$ are hyperparameters tuned by a cross-validation process.
We choose to use the Wasserstein distance as the GAN loss function (WGAN)~\cite{arjovsky2017wasserstein}, which improves the training stability and provides better reconstruction compared to the classic GAN loss we used in our preliminary work~\cite{shitrit2017accelerated}.

@@

The training data consists of 500 3D brain MRI (T1-weighted) scans of different patients from the IXI dataset.\footnote{\url{http://brain-development.org/ixi-dataset/}} The data was acquired by three MR machines, Philips 1.5T, 3T and GE 1.5T using different imaging protocols.
From each 3D volume we extracted 93 2D saggital slices. All slices were interpolated to $256\times256$ pixels. We used $37.2k$ $(80\%)$ 2D slices for training and $9.3k$ $(20\%)$ for testing (100 3D volumes).
In order to create k-space images for training, inverse orthonormal
2D FFT was applied to the fully-sampled MR images. 
Unlike our preliminary work~\cite{shitrit2017accelerated}, which employed a 1D Cartesian sampling mask along the phase axis; here, we perform k-space sampling using a 2D random Gaussian mask. By this, we exploit the sparsity in both phase and frequency dimensions and are therefore able to accommodate higher sampling factors (Figure~\ref{fig:sampling_masks}). MRI reconstruction is performed using 40\%, 25\%, and 16.6\% of the original k-space data.
%Data augmentation is created by random offsets of raw images (prior to the IFFT) and k-space flipping. This leads us to reconstruction of the MR image from 40\%, 25\% and 16.6\% of the original k-space data.

The generator is composed of $5$ blocks of CONV-BatchNorm-ReLU, with output channels $16,32,64,32,8,$ respectively (see Figure~\ref{fig:architectures}). The last CONV layer is composed of two output channels for real and imaginary components. The discriminator is composed of $4$ blocks of CONV-Pool-BatchNorm-LReLU with output channels $8,16,32,16$ and one fully-connected layer. All CONV layers' kernel size is $3\times3$. All weights were initialized using Xavier \cite{glorot2010understanding}. We used the RMSprop~\cite{tieleman2012lecture} solver with a fixed learning rate of $5\mathrm{e}{-6}$ and set $k_{d}$ to 1.

We compared the proposed method reconstructions to those obtained by using a conventional compressed sensing method-CS-MRI \cite{lustig2007sparse} and zero-filling. In addition, we tested two other networks. To assess the use of an adversarial loss, we trained a k-space generator, with the same architecture as the proposed method, with only an L2 loss (CNN-L2). We also implemented the method proposed in~\cite{hyun2017deep} and constructed a network IM-CNN-L2, which is applied directly to the image space rather than the k-space. Specifically, as in~\cite{hyun2017deep}, we used the U-net architecture~\cite{ronneberger2015u} and applied an FFT on each of its output images to get an estimated k-space. The complete k-space is then recovered by integrating the estimated samples with the sampled raw data according to the sampling mask. The same sampling masks were used for all methods.

To demonstrate the clinical usability of the reconstructed images we performed an extended validation, composed of anatomical quantitative measurements. In addition to the commonly used PSNR and Structure Similarity Index (SSIM)~\cite{wang2004image} metrics, we performed tissue segmentation and compared the results to those obtained by using the original fully-sampled images. The segmentation compatibility is measured in terms of Dice scores and Hausdorff distances.

@@

\subsection{PSNR and SSIM}
PSNR and SSIM are the most commonly used measures for evaluation of image reconstruction quality. PSNR measures the Mean Squared Error (MSE) between the fully-sampled MR image and the reconstructed image.
Quantitative evaluation is presented in~Table~\ref{tbl:PSNR_NO_MASK} and a graphical visualization in ~Figure~\ref{fig:error_psnr_errorbar}. The PSNR is calculated for the entire image without masking. Note that the proposed method outperforms the others. To support this, we also calculated the two-sample t-Test for the table's scores. The differences are statistically significant (p-values~$<0.005$) for all pairwise comparisons between the results of our method and the others.

@@

An additional validation of our model is performed by calculating the PSNR measure on different brain tissues. We applied the commonly used FAST~\cite{zhang2001segmentation} image segmentation algorithm on the original fully-sampled MR image and then used it for calculating the masked-PSNR on gray matter (GM), white matter (WM), and the Cerebrospinal Fluid (CSF). Results are presented in~Table~\ref{tbl:PSNR_WITH_MASK}. 
The differences are statistically significant with p-values~$<0.0005$ for CS-MRI and zero-filled for sampling factors 2.5 and 4, and also for IM-CNN-L2 with p-values~$<0.0001$ for sampling factor 6.

@@

We next used SSIM, which is a full reference metric, and is considered a perception-based model, to measure the similarity between the original fully-sampled images and the reconstructed ones.
It quantifies the degradation caused by sampling artifacts based on pixels' inter-dependencies in the image. 
%The SSIM between images $I_{\mbox{\tiny{REF}}}$ and $I_{\mbox{\tiny{REC}}}$ is calculated on a multiples sliding windows with size of $N\times N$ as follows:
%\begin{equation}
%\begin{array}{cc}
%  SSIM(I_{\mbox{\tiny{REF}}},I_{\mbox{\tiny{REC}}}) = \frac{(2\mu_{I_{\mbox{\tiny{REF}}}}\mu_{I_{\mbox{\tiny{REC}}}} + c_1) (2\sigma_{I_{\mbox{\tiny{REF,REC}}}} + c_2)}{(\mu_{I_{\mbox{\tiny{REF}}}}^2 + \mu_{I_{\mbox{\tiny{REC}}}}^2 + c_1) (\sigma_{I_{\mbox{\tiny{REF}}}}^2) + \sigma_{I_{\mbox{\tiny{REC}}}}^2) + c_2} \\
% second = 2\\
% ohad
%\end{array}
%\end{equation}
%Let $\mu_1,\mu_2$ denote the brain boundaries ext
SSIM scores for different sampling factors, for all methods compared, are presented in~Table~\ref{tbl:SSIM_NO_MASK}. The advantage of our method is clearly shown. All pairwise comparisons between the results obtained by our method and the others are statistically significant (p-values~$<0.0001$) for all sampling factors.

@@

Let $C_1,C_2\in R^2$ denote the brain boundaries extracted by the BET algorithm, using the original and the reconstructed scans, respectively. The MHD between $C_1$ and $C_2$ is:

\begin{equation}
\begin{array}{cc}
MHD(C_1,C_2) = 0.5 \left[{\frac{1}{|C_1|} \sum_{\sigma_{1}\in C_1}^{}d(\sigma_1,C_2) + \frac{1}{|C_2|} \sum_{\sigma_{2}\in C_2}^{}d(\sigma_2,C_1)}\right] \\
d(\sigma_1,C_2) = \underset{\sigma_{2}\in C_{2}}{\min}||\sigma_1-\sigma_2|| \\
d(\sigma_2,C_1) = \underset{\sigma_{1}\in C_{1}}{\min}||\sigma_2-\sigma_1||
\end{array}
\end{equation}
where $\sigma_1,\sigma_2$ are points on the contours $C_1,C_2$, respectively (Figure~\ref{fig:brain_extraction}.c) and $|C_i|,~i=1,2$, are the respective cardinalities. MHD values for all tested methods and sampling ratios are presented in~Table~\ref{tbl:MHD}. The MHD results obtained by our methods are better than all others and are significant (p-values~$<0.0001$) for CS-MRI and zero-filled.

@@

	\caption{\textcolor{black}{\footnotesize{}{}Segmentation Dice scores for different sampling ratios}{\footnotesize{}\label{tbl:DICE}}}

\end{table}
MR reconstruction results for factor 6~(16.66\%) as well as their respective brain segmentation results are visually presented in Figure~\ref{fig:example_factor_6} for a single representative scan. The left most column corresponds to the original image and columns 2-5 correspond to the reconstructed images using Zero-filled, CS-MRI, IM-CNN-L2, and the proposed method, respectively. The first row presents a $2D$ slice of the original and the reconstructed MRIs. The second row presents the zoom-in of brain regions in the red frames shown in the first row. The third row presents the corresponding brain extraction. The fourth row presents the automatic tissue segmentation where, dark blue is used for WM, blue for GM and light blue for CSF. Note that MRIs reconstructed using the suggested adversarial method have stronger contrast and no significant aliasing or artifacts. In addition, the brain extraction and tissue segmentation results are more compatible with those obtained from the original fully-sampled MRIs. Additional examples of different cases, using sampling factors of 4~(25\%), and 2.5~(40\%) are shown in Figures~\ref{fig:example_factor_4},~\ref{fig:example_factor_2.5}, respectively.
